### 2018-11-14 version3supportFiles/PLA/highlights.tex
Added Kishor's changes and now working on changing the format to PLA

\documentclass[12pt]{article}
\addtolength{\topmargin}{-2cm}
\addtolength{\textheight}{3cm}
\addtolength{\oddsidemargin}{-1.5cm}
\addtolength{\textwidth}{2cm}
\usepackage{graphicx}
\pagestyle{empty}
\begin{document}
\thispagestyle{empty}
\centerline{\large \bf Highlights for Review}
\vspace*12pt
In this work we propose a new (to the best of our knowledge)
non-contextual hidden variable model which helps clarify the
limitations of a well-established notion in the foundations
of quantum mechanics, that of contextuality. Contextual
hidden variable models on the other hand are already known
but these are in stark contrast with our model.
Understanding the notion of contextuality and its
implications better is therefore likely to stimulate further
research. Our work contributes towards this older direction
which has witnessed a renewed interest.
\end{document}